A geo-referenced micro-data set of real estate listings for Spain's three largest cities

### 2023-05-17 editorial-materials/EPB-2022-0116.R1_response_to_reviewer.Rmd

```{r setup,  cache=FALSE, warning = FALSE, include=FALSE, message = FALSE}
library(idealista18)
library(tidyverse)

set.seed(1234)
```
-->

Referee: 1

>Comments to the Author

>Dear Authors, Thank you for the opportunity to review this paper. I would appreciate it if you would highlight the changed text in the next submission. Otherwise it is difficult to identify which parts have changed. The unknown quality of this one year randomised data is the main issue for potential users and I repeat my suggestion that this data should be validated for the benefit potential readers and users. Below is a list of comments and questions which  need to be addressed.

Thank you for your constructive comments. Below we describe how we improved the paper in response.

>1. Please clearly state the nature of the data in the abstract. It is worthwhile to mention the relocated and randomised dataset in the abstract. This house price dataset only has 189,923 records, which could be not really be described as a big dataset as you describe it in the abstract. Except the authors could provide an evidence/reference on why this is big dataset.

@@

Thank you for highlighting this. We have carefully proof-read the document for typos and other errors.

---

>3. Please use part of your answer to my previous fourth question in the manuscript. Your answer is quite useful for your future users, especially those researchers who are not familiar with house price data in Spain. I quote your previous answer below:

>“idealista covers fairly well all segments in the Spanish market, both individual and professional advertisers. This dataset comprehends information about listing prices, thus it contains the market situation from the offer. Nevertheless, actual sale prices information is not publicly available as it is information that can only be accessed by paying high fees to Colegio de Registradores. However listing prices reflect quite well the (transaction) reality of real estate markets using the offer perspective, they keep strong correlation between idealista and transaction prices can be established see Banco de España. Even though a great extent of correlation does exist between official transactional and asking records, both databases can be taken as complementary and are of great interest when studying asking-transaction price gaps or the relation between listing site demand variables (ie. ad contacts or ad views) and price gaps. The main real estate portals in Spain are idealista and Fotocasa, further away are Habitaclia and Milanuncios, the latter focusing almost exclusively on private individuals. In September 2021, according to data from Similarweb (a site specialized in sites’ traffic volume comparison), in Spain, there were a total of 103 million page views on real estate portals, out of page views on real estate portals, of which the four main portals, idealista, Fotocasa, Habitaclia, and Pisos.com, in that order, accounted for 94% of the traffic. Another relevant feature of this sector is that it is highly concentrated, idealista being the leader by far with 58.6 monthly million visits (57% of the total traffic) versus its immediate competitor with 19.9 million visits (19.3 % of total traffic). In terms of the evolution over time of interest in the content of each portal, based on data from Google Trends, the leadership of idealista Trends, the leadership of idealist”

@@

**Response:**

Thank you for this suggestion. We have adapted this answer to expand the text in the section "Data Description".

>4. For data description, it is worth showing solid evidence of the quality of this dataset. For example, what is the data coverage of this data compared to all the residential properties in the city?  Listing the advertised volume is not very helpful for users who wish to understand the quality.

To better place in context the coverage of the {idealista18} dataset, Table \ref{tab:transactions} now shows the number of listings with respect to the total residential stock in each city (in 2018). As seen in the table, the number of listings ranges between 6.1% of the total number of properties (in Madrid) and 8.1% (in Valencia). Information from Instituto Nacional de Estadistica^[\url{https://www.ine.es/jaxiT3/Tabla.htm?t=6150&L=1}] for 2018 also shows that the number of transactions in 2018 are equivalent to 81.3\% of idealista listings in Barcelona, 80.8\% in Madrid, and 91.1\% in Valencia. While it is not possible to link actual transactions to idealista listings, this gives a quantitative sense of the coverage of the data package.

@@

---

>5. Given the asking price is a resized version of the original asking price with a random percentage between -2.5% and +2.5%, please elaborate on how this influences your quantitative research results and findings. Please provide critical evidence to explain how this randomised price will not influence academic research. For example, how does the house price variation at local level change after the price is randomised. To what degree will this influence a typical hedonic house price research approach? I am concerned about whether this randomised data could bias conclusions.

**Response:**

We appreciate the opportunity to delve more deeply into this issue, as it may be of interest to users of the data package. The short version of this story is that adding a small amount of random noise to the prices ($\pm 2.5\%$) does not introduce bias but does increase the variance of the masked variable. This can be shown with some standard algebra of random variables.

Now the longer story.

The masked prices ($P$) in the dataset are calculated as the sum of two random variables, the original raw prices ($RP$) and a random masking variable ($\eta$). The masking variable is generated as follows:
$$
\eta = RP\cdot\epsilon
$$

The masking variable uses values drawn at random from the uniform distribution with parameters $b = 0.025$ and $a = -0.025$:
$$
f(\epsilon) = 
\begin{cases}
\frac{1}{b - a} & \text{for } a\le \epsilon \le b\\
0 & \text{otherwise}
\end{cases}
$$

The effect of $\epsilon$ is to produce a random percentage of the raw price ($\pm 2.5\%$). The expectation of $\epsilon$ given these parameters is:
$$
E[\epsilon] = \frac{a+b}{2} = \frac{-0.025 + 0.025}{2} = 0
$$

\noindent and the variance is:
$$
\mathrm{V}[\epsilon]=\frac{(b - a)^2}{12} = \frac{25}{120000}
$$

The masked prices $P$ are the raw prices plus the masking variable:
$$
P = RP + RP\cdot\epsilon = RP + \eta
$$

Since the mean of $\epsilon$ is zero, and $RP$ and $\epsilon$ are independent, it turns out that the mean of the masking variable is zero:
$$
\mathrm{E}[\eta] = \mathrm{E}[RP\cdot\epsilon] = \mathrm{E}[RP] \cdot \mathrm{E}[\epsilon] = 0
$$

It follows that the masked prices are unbiased since their expectation is identical to the expectation of the raw prices:
$$
\mathrm{E}[P] = \mathrm{E}[RP] + \mathrm{E}[\eta] = \mathrm{E}[RP]
$$

Now let us examine the covariance. Again, since $RP$ and $\epsilon$ are independent, the variance of the masking variable $\eta$ is:
$$
\mathrm{V}[\eta] = \mathrm{V}[RP\cdot\epsilon] = \mathrm{V}[RP]\cdot\mathrm{V}[\epsilon] + \mathrm{V}[RP]\cdot(\mathrm{E}[\epsilon])^2 + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2  
$$

Considering that $\mathrm{E}[\epsilon] = 0$, then:
$$
\mathrm{V}[\eta] = \mathrm{V}[RP]\cdot\mathrm{V}[\epsilon] + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2 = \mathrm{V}[\epsilon]\cdot\big(\mathrm{V}[RP] + (\mathrm{E}[RP])^2\big) = \frac{25}{120000}\big(\mathrm{V}[RP] + (\mathrm{E}[RP])^2\big)  
$$

Accordingly, the variance of the masked prices is:
$$
\mathrm{V}[P] = \mathrm{V}[RP + \eta] = \mathrm{V}[RP] + \mathrm{V}[\eta] + 2\mathrm{Cov}[RP,\eta] 
$$

If we recall that $\mathrm{E}[\eta] = 0$ and $\mathrm{E}[\epsilon] = 0$, the covariance of $RP$ and $\eta$ can be shown to be zero:
$$
\mathrm{Cov}[RP,\eta] = \mathrm{E}[RP\cdot \eta] - \mathrm{E}[RP]\cdot \mathrm{E}[\eta] = \mathrm{E}[RP\cdot \eta] = \mathrm{E}[RP^2\cdot \epsilon] = \mathrm{E}[RP^2]\cdot \mathrm{E}[\epsilon] = 0
$$

From the above we can conclude that the variance of the masked prices is:
$$
\mathrm{V}[P] = \mathrm{V}[RP] + \mathrm{V}[\eta] = \mathrm{V}[RP] + \frac{25}{120000}\big(\mathrm{V}[RP] + (\mathrm{E}[RP])^2\big) = \frac{120025}{120000}\cdot\mathrm{V}[RP] + \frac{25}{120000}\cdot(\mathrm{E}[RP])^2
$$

\noindent which is $0.0208%$ greater than the variance of the raw prices, plus $0.0208%$ of the square of the mean of the raw prices.

What does this mean for users of the data? The masked prices are an unbiased version of the raw prices, but they have a somewhat larger variance than the original (unreported) prices.

@@

---

>6. In figure 1, is it possible to standardise this data rather than showing only the number of dwelling records in your house price data? I guess the city centre always has high advertised volumes but it also has a high density of dwellings. It is worthwhile to show values standardised by the number of dwellings.

After this comment we included a new figure in the paper to display the percentage of listings relative to the total number of dwellings by district (see Figure \ref{fig:districts} below).

```{r, echo = FALSE,message=FALSE, fig.height=8, fig.cap="\\label{fig:districts}Percentage of listings relative to total number of dwellings by district. Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
library(dplyr)
library(ggplot2)
library(glue)
library(gridExtra)
library(idealista18)
library(sf)

@@

# Barcelona
Barcelona_prop <- properties_by_district |>
  filter(CITY == "Barcelona") |>
  mutate(proportion = N/N_CADASTRE)

q <- quantile(Barcelona_prop$proportion)
pq <- round(q * 100, 2)

Barcelona_prop <- Barcelona_prop |>
  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))

plot4 <- ggplot(data = Barcelona_prop) +
  geom_sf(aes(fill = Quantile),
          color = NA) +
  geom_sf(data = Barcelona_Polygons,
          fill = NA) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank())

@@

# Madrid
Madrid_prop <- properties_by_district |>
  filter(CITY == "Madrid") |>
  mutate(proportion = N/N_CADASTRE)

q <- quantile(Madrid_prop$proportion)
pq <- round(q * 100, 2)

Madrid_prop <- Madrid_prop |>
  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))

plot5 <- ggplot(data = Madrid_prop) +
  geom_sf(aes(fill = Quantile),
          color = NA,
          size=.2) +
    geom_sf(data = Madrid_Polygons,
          fill = NA) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank())

@@

# Valencia
Valencia_prop <- properties_by_district |>
  filter(CITY == "Valencia") |>
  mutate(proportion = N/N_CADASTRE)

q <- quantile(Valencia_prop$proportion)
pq <- round(q * 100, 2)

Valencia_prop <- Valencia_prop |>
  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))

plot6 <- ggplot(data = Valencia_prop) +
  geom_sf(aes(fill = Quantile),
          color = NA,
          size=.2) +
    geom_sf(data = Valencia_Polygons,
          fill = NA) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank())

@@

# Arrange plots
grid.arrange(plot4, plot5, plot6, 
             ncol = 1, 
             nrow = 3)
```

---

Thank you again for your generous feedback. We hope that our responses will prove satisfactory and you will agree that the paper makes a worthwhile contribution to growing catalog of datasets to support research on real estate markets.

\newpage

## References

### 2023-05-17 EPB/main_EPB.Rmd
Further work in preparation for resubmission

The data set has been supplied directly by Idealista, and therefore is clean and free of download errors. However, in order to be able to share publicly the data we have slightly modified some variables by applying some random noise which will not bias the main results derived from its usage.

@@

The data set is distributed as an `R` package, named {idealista18}, which can be accessed from a public Github repository\footnote{\url{https://github.com/paezha/idealista18}}. The open data product is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}. For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.

@@

The open data product {idealista18} is an `R` package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia. For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the package.

The data provider (idealista.com) is a leading real estate portal in Spain, on par with Fotocasa. Smaller portals include Habitaclia and Milanuncios, the latter focusing almost exclusively on individual (i.e., non-professional) advertisers. In September 2021, according to data from Similarweb (a site specialized in sites' traffic volume comparison), there were a total of 103 million page views on real estate portals in Spain. Of this, 94% of traffic was concentrated on four portals, of which idealista was the most important, followed by Fotocasa, Habitaclia, and Pisos.com. Traffic is also highly concentrated, and idealista.com was the the leader by far with 58.6 monthly million visits (57% of the total traffic) compared to its closest competitor with 19.9 million visits (19.3 % of total traffic).

@@

plot4 <- ggplot(data = Barcelona_prop) +
  geom_sf(aes(fill = Quantile),
          color = "black",
          size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

plot5 <- ggplot(data = Madrid_prop) +
  geom_sf(aes(fill = Quantile),
          color = "black",
          size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

plot6 <- ggplot(data = Valencia_prop) +
  geom_sf(aes(fill = Quantile),
          color = "black",
          size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

>>>>

The data set has been supplied directly by Idealista, and therefore is clean and free of download errors. However, to comply with data legislation, to share the data we have slightly modified the prices by applying a small amount of random noise which will not bias the main results derived from its usage.

@@

The data set is distributed as an `R` package, named {idealista18}, which can be accessed from a public Github repository\footnote{\url{https://github.com/paezha/idealista18}}. The open data product is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}. For transparency, we also share the randomization process applied to the original data in the aforementioned Github repository.

@@

The open data product {idealista18} is an `R` package composed of ten objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia. For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the package. There is in addition a data object with the number of dwellings by district (a collection of neighborhoods), according to the Spanish cadastre.

The data provider (idealista.com) is a leading real estate portal in Spain, on par with its nearest competitor Fotocasa. Smaller listing portals include Habitaclia and Milanuncios, the latter focusing almost exclusively on individual (i.e., non-professional) advertisers. In September 2021, according to data from Similarweb (a site specialized in sites' traffic volume comparison), there were a total of 103 million page views on real estate portals in Spain. Of this, 94% of traffic was concentrated on four portals, of which idealista was the most important, followed by Fotocasa, Habitaclia, and Pisos.com. Traffic is also highly concentrated, and idealista.com was the the leader by far with 58.6 monthly million visits (57% of the total traffic) compared to its closest competitor with 19.9 million visits (19.3 % of total traffic).

@@

plot4 <- ggplot(data = Barcelona_prop) +
  geom_sf(aes(fill = Quantile),
          color = NA) +
  geom_sf(data = Barcelona_Polygons,
          fill = NA) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

plot5 <- ggplot(data = Madrid_prop) +
  geom_sf(aes(fill = Quantile),
          color = NA,
          size=.2) +
    geom_sf(data = Madrid_Polygons,
          fill = NA) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

plot6 <- ggplot(data = Valencia_prop) +
  geom_sf(aes(fill = Quantile),
          color = NA,
          size=.2) +
    geom_sf(data = Valencia_Polygons,
          fill = NA) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

### 2023-05-17 EPB/main_EPB.tex
Further work in preparation for resubmission

The data set is distributed as an \texttt{R} package, named
\{idealista18\}, which can be accessed from a public Github
repository\footnote{\url{https://github.com/paezha/idealista18}}. The
open data product is made available under the
\href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
For transparency, we also share the randomisation process applied to the
original data in the aforementioned Github repository.

@@

The open data product \{idealista18\} is an \texttt{R} package composed
of nine objects, three objects for each of the three main Spanish
cities: Barcelona, Madrid, and Valencia. For each city, dwelling
listings, neighborhood polygons, and a set of points of interest have
been included in the package.

The data provider (idealista.com) is a leading real estate portal in
Spain, on par with Fotocasa. Smaller portals include Habitaclia and
Milanuncios, the latter focusing almost exclusively on individual (i.e.,
non-professional) advertisers. In September 2021, according to data from
Similarweb (a site specialized in sites' traffic volume comparison),
there were a total of 103 million page views on real estate portals in
Spain. Of this, 94\% of traffic was concentrated on four portals, of
which idealista was the most important, followed by Fotocasa,
Habitaclia, and Pisos.com. Traffic is also highly concentrated, and
idealista.com was the the leader by far with 58.6 monthly million visits
(57\% of the total traffic) compared to its closest competitor with 19.9
million visits (19.3 \% of total traffic).

@@

\begin{figure}
\centering
\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:all-polygons}Quantile maps of the number of
dwellings in each neighborhood. Boundary for Barcelona (Left), Madrid
(Center), and Valencia (Right).}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:districts}Quantile maps of the percentage of
listings relative to total number of dwellings by district. Boundary for
Barcelona (Left), Madrid (Center), and Valencia (Right).}
\end{figure}

@@

In terms of the asking prices, the original values were obfuscated with
the addition or subtraction of a random percentage of their original
values, ranging from -2.5\% to +2.5\%. Since asking prices are usually
multiples of 1,000, after the first price modification, the prices were
aligned to multiples of 1,000.

>>>>

The data set is distributed as an \texttt{R} package, named
\{idealista18\}, which can be accessed from a public Github
repository\footnote{\url{https://github.com/paezha/idealista18}}. The
open data product is made available under the
\href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
For transparency, we also share the randomization process applied to the
original data in the aforementioned Github repository.

@@

The open data product \{idealista18\} is an \texttt{R} package composed
of ten objects, three objects for each of the three main Spanish cities:
Barcelona, Madrid, and Valencia. For each city, dwelling listings,
neighborhood polygons, and a set of points of interest have been
included in the package. There is in addition a data object with the
number of dwellings by district (a collection of neighborhoods),
according to the Spanish cadastre.

The data provider (idealista.com) is a leading real estate portal in
Spain, on par with its nearest competitor Fotocasa. Smaller listing
portals include Habitaclia and Milanuncios, the latter focusing almost
exclusively on individual (i.e., non-professional) advertisers. In
September 2021, according to data from Similarweb (a site specialized in
sites' traffic volume comparison), there were a total of 103 million
page views on real estate portals in Spain. Of this, 94\% of traffic was
concentrated on four portals, of which idealista was the most important,
followed by Fotocasa, Habitaclia, and Pisos.com. Traffic is also highly
concentrated, and idealista.com was the the leader by far with 58.6
monthly million visits (57\% of the total traffic) compared to its
closest competitor with 19.9 million visits (19.3 \% of total traffic).

@@

\begin{figure}
\centering
\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:all-polygons}Number of listings in each
neighborhood. Boundary for Barcelona (Top), Madrid (Center), and
Valencia (Bottom).}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:districts}Percentage of listings relative to total
number of dwellings by district. Boundary for Barcelona (Top), Madrid
(Center), and Valencia (Bottom).}
\end{figure}

@@

In terms of the asking prices, the original values were obfuscated with
the addition or subtraction of a random percentage of their original
values, ranging from \(-2.5%
\) to \(+2.5%
\). Since asking prices are usually multiples of 1,000, after the first
price modification, the prices were aligned to multiples of 1,000.
